LME-MPM applied to quasi-brittle fracture.

### 2020-05-27 _region_.tex
Correcciones sobre PNA+DM 20-5-2020
Faltan cosas por modificar sobre lo corregido

As noticed in \cite{Zhang_EE_2020}, the EigenMPM have two important
shortcomings. The first is the presence of stress instabilities even
with \acrshort{gimp} shape functions. To overcome it, \acrshort{lme}
approximants are introduced as an
alternative to the existing interpolation techniques. Its
impressive performance mitigating spurious stress oscillations
\cite{Wobbes2020} under the \acrshort{mpm} framework help to enhance
notoriously the quality of the results as we can see in Section
\ref{sec:3.1}. Where both interpolation techniques
are compared by carry out an eigenerosion simulation. The second
limitation is concerning its inability to simulate properly quasi-brittle
fracture \cite{Navas_2018_ES}. It can be solved through the
eigensoftening algorithm described in Section \ref{sec:2.3}. A proof
of it is exposed in Section \ref{sec:3.3}, where experimental results
of a drop-weight impact test are compared with those produced by
eigensoftening computations.

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}

This problem here presented is devoted to compare how much \acrshort{lme}
approximants are able to improve the result \textit{versus} standard
linear interpolation. It consists in a square plate of size $H = 1$ containing
an initial edge crack of length $0.25 \cdot H$ loaded in a pure mode I
by displacement control on the outer flanks of the plate, Figure
\ref{fig:geometry-cracked-panel-mode-I}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model consider for numerical experiment is a
linear-elastic Hookean material, where Young's modulus \gls{E} = 1.06, the
Poisson's ratio \gls{nu} = 0.333, and critical energy-release rate
\gls{Gf} = 0.0001. This set of parameters was taken from \cite{Pandolfi_2012}. In
\acrshort{mpm}, two different discretization are required. In one hand
a cartesian grid of nodes is consider with a nodal spacing value of
0.025. On the other hand, the plate will be modeled with a initial
layout of four particles per element and occupying the Gauss quadrature
positions.\\

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in both loading and fracture stages
in the reaction-displacements curve when linear interpolation
technique is used. In contrast, \acrshort{lme} simulation does not
exhibit these spurious oscillations and remains linear. Besides that,
linear interpolation produces stiffer results than
\acrshort{lme}. This can be attributed to imprecision in the stress
field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
On the other hand, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case. Although this behavior does not affect significantly to the peak
stress value in a hookean material \cite{Zhang_EE_2020}, if
hyperelastic materials are consider for future simulations,
inaccuracies in the stress field could affect dramatically to the
final result. On the other hand \acrshort{lme} produces soft stress
field evolution even after breaking.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated, Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver.

\subsection{Drop-weight impact test}
\label{sec:3.2}

Once the accuracy of the \acrshort{lme} approximants have been proven
in Section \ref{sec:3.1}, henceforth will be adopted without comparing
with linear interpolation. Therefore this section is devoted to proof
the accuracy of the eigensoftening algorithm to simulate the behavior
of quasi-brittle materials under dynamic loading. An example of this
load case is the three-point bending test on a notched beam is
conducted under impact loading. This kinds of experiments are
challenging to reproduce numerically with an explicit solver since
fracture occurs in a period of milliseconds and high strength
materials are involved. Accordingly, really small time steps are
required to achieve a simulation numerically stable. Additionally,
this kind of simulations traditionally require the implementation of
contact algorithm to reproduce the hammer impact in the beam. In the
\acrshort{mpm}, the implementation of this feature can be bypassed to
obtain preliminary results. Nevertheless, this kind of algorithm could
be implemented to get ever more accurate solutions. Although this is
out of the scope of this document, future research of authors will overcome this
limitation.\\

As in the study of Zhang {\it et al.} \cite{Zhang_2009,Zhang_2010a},
an impact hammer of 120.6 kg was employed to drop it at an impact
speed of 881 mm/s. The beam dimensions where 100 mm x 100 mm (B x D)
in cross section, and 420 mm in total lenght (L). The initial
notch-depth ratio was approximately 0.5, and the span, S, was fixed at
300 mm during the tests, see Figure
\ref{fig:geometry-drop-weight-impact-test}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./Figure-impact-test}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-drop-weight-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material adopted for the simulation was characterised by Navas
{\it et al.}~\cite{Navas_2017_ES}, and the material properties, such
as the material density, \gls{rho}, the compressive strenght, \gls{fc},
the tensile strenght \gls{ft}, the specific fracture energy, \gls{Gf},
and the elastic modulus \gls{E} are given in table \ref{tab:concrete-properties}.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented. Several levels of discretization were required to assess
the objectiveness of the obtained results. The results here exposed
are from a discretization of 30616 material points and 6297 nodes. To
achieve better results, an unestructured grid layout has been adopted
focusing the minimum nodal spacing, 0.47 mm, in the middle of the beam and in
the impact surface of the hammer. For this simulation, \gls{gamma-lme} was
fixed to 6.\\

First the reaction and impact forces are validated against their
experimental counterparts. Since the impact forces applied by the
hammer and the reaction forces at the two supports were experimentally
measured, they are compared with numerical ones in Figure
\ref{fig:Reactions-Forces-impact-test}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-impact-test-Forces-Time}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-Forces-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Note that the general trend of both forces are correctly
captured. Observe that in experimental solution the delay between
impact peak and reactions peak does not agree with the numerical
ones. On the other hand, if the delay is computed analytically through
concrete wave speed is much closer to the numerical results. A
feasible explanation to this unexpected result can due the composite
nature of the concrete. Another discrepancy is in the much higher
values of maximum impact and reaction forces in numerical than in
experimental. In this case, this is a consequence of over-rigidity
in the supports. Therefore, this results can be improved in future research
by a suitable contact algorithm.\\

And finally horizontal stress distributions and crack propagation are
extracted, see Figure \ref{fig:Stress-vs-damage-impact-test}. Stress
evolution agrees with the results obtained by \cite{Navas_2017_ES}
under the \acrshort{otm} framework. Two additional considerations about
this results are concerning the interpolation technique. In one hand,
in the opposite to the \acrshort{ugimp} shape function, \acrshort{lme}
allows unstructured grids. This is a remarkable benefit for this kind
of simulations since the majority of the computational effort is
clearly localised in a tiny region of the domain. And in the other
hand there is the possibility to reduce the shape function support to avoid
a material point in one side of the crack is affected by those in the
other side of the crack. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 0.27 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0060}}
\subfigure[t = 0.36 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0080}}
\subfigure[t = 0.45 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0100}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-vs-damage-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:4}
Based on the EigenMPM approach to fracture proposed by Zhang {\it et al.}
\cite{Zhang_EE_2020}, we have proposed two addition improvements which
solves the shortcomings presented in the original concept. First by
bring to the \acrshort{mpm} framework the concept of eigensoftening,
which allow to consider gradual failure process of each material
point. This concept can be compared with a cohesive law in the context
of cohesive fracture. Furthermore, more elaborated damage curves can
be designed to describe the failure process in more sophisticated
materials. In this direction, an interesting future improvement of the
algorithm would be the ability to simulate of crack patterns in
composite material such carbon fiber. This leads a challenging
research topic. Secondly, \acrshort{lme} is employed to remove from
the solution those wiggles due to inaccuracies in the stress
integration. This interpolation technique could help to improve
results in fracture propagation by adapting it through the
deformation gradient. \textit{A priori} this will enhance the already
impressive localisation properties of eigensoftening and eigenerosion
algorithms.

@@

\end{document}
\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-1135) }

>>>>

\DM{The proposed approach to overcome the two main shortcomings of EighenMPM \cite{Zhang_EE_2020} has been evaluated with "two benchmark tests"¿?. } \DM{DELETE: As noticed in \cite{Zhang_EE_2020}, the EigenMPM have two important
shortcomings}. The first one deals with the presence of stress instabilities provoked by the spatial discretization, appearing even
when the \acrshort{gimp} shape functions are employed. To overcome it, \acrshort{lme}
approximants are introduced as an
alternative to the existing interpolation techniques. Its
performance in mitigating spurious stress oscillations under the \acrshort{mpm}~\cite{Wobbes2020,Molinos2020} framework helps to enhance
notoriously the quality of the results as it is observed in Section
\ref{sec:3.1}, where both interpolation techniques
are compared by carrying out an eigenerosion simulation. The second
limitation is concerned to its inability to simulate properly quasi-brittle
fracture \cite{Navas_2018_ES}. It can be solved through the
eigensoftening algorithm described in Section \ref{sec:2.3}.\MMP{A proof
of it is exposed in Sections \ref{sec:3.2} \ref{sec:3.3}, where experimental results
are compared with experimental ones as well as eigensoftening
computations. In the first case comparing with a semicircular bending
test and in the second case with a drop-weight impact test.}

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented hereinafter is to assess the
capability of the \acrshort{lme} approximants to improve the result
\textit{versus} the standard linear interpolation. The application
consists of a square plate of size $H = 1$ containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\PNA{UNIDADES}

\PNA{NOMBRAR ALGO A LOS CASOS ANALITICOS O DE MICHAEL ORTIZ}

The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \gls{E} = 1.06, the
Poisson's ratio \gls{nu} = 0.333, and critical energy-release rate
\gls{Gf} = 0.0001. This set of parameters has been previously employed by Pandolfi~\textit{et al.}~\cite{Pandolfi_2012}. In
\acrshort{mpm}, two different discretization are required. On the one hand,
a cartesian grid of nodes is considered with a nodal spacing value of
0.025. On the other hand, the plate will be modeled with a initial
layout of four particles per element and occupying the Gauss quadrature
positions.\\

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in the reaction-displacements curve, in both loading and post-failure stages,
when linear interpolation technique is used. In contrast, \acrshort{lme} simulation does not
exhibit these spurious oscillations and remains linear until the failure. About the strength,
linear interpolation produces stiffer results than
\acrshort{lme}. This can be attributed to imprecision in the stress
field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. \PNA{Although, apparently, the differences in the obtained peaks are not extremely significant, it can be owing to the aforementioned hookean material employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive models will be employed, severe
%if hyperelastic materials are consider for future simulations,
inaccuracies may appear in the stress field that could affect dramatically to the
final result. Focusing on the post failure behavior, it can be seen how \acrshort{lme} produces soft stress
field evolution even after breaking.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: \PNA{once the
  energy is release dinamically by the crack, the waves propagate
  along the domain.}

\subsection{Semicircular bending test}
\label{sec:3.2}

% The second application to be considered is a series of semicircular
% bending tests on Johnstone, which have been carried out by Lim et al.95 The aim of this example is to examine its predictive capability of the SPH model under a wide range
% of mixed-mode loading conditions. The geometry and boundary conditions of the test in the experiment are shown in
% Figure 9A. Similar with 3-point bending test, the semicircular specimens with a radius of 47.5 mm and a thickness of
% 20 mm are supported by two rollers at a span of 47.5 mm and loaded by another roller on its mid-span. Notches with a
% length of 16.6 mm are created with different notch angles varying from 0 to 60 with respect to the vertical axis to
% investigate the influences of the notch angles on the peak load (P) and the fracture toughness (KI,KII), which are calcu-
% lated as follows:

% where KI and KII are respectively mode I and mode II fracture toughness; r is the specimen radius; t is the specimen
% thickness; a is the notch length; YI and YII are mode I and mode II normalised stress intensity factor. As the values of YI
% and YII can be obtained from the data by Lim at al.6 KI and KII can be calculated based on the above equation once the
% peak load (P) is known. The material properties of Johnstone are listed in Table 1. The friction angle and dilation angle
% are respectively taken as 33 and 4  for average quality rock masses in Hoek and Brown96 for the following examples.
% Figure 9 shows the corresponding SPH configuration of the semicircular bending test with a notch angle of 30 . In
% the tests, a particle discretisation of 0.95 mm is used, resulting in a total number of 3995 SPH particles for generating
% the specimen. To create accurate notch angles using reasonable particle numbers, the initial SPH arrangement in this
% test is different from the traditional one such as the 3-point bending test in Section 5.1. The initial location of each SPH
% particle in semicircular bending tests is related to the notch angle as shown in Figure 9B. Similar to the 3-point bending
% tests, three layers of boundary particles are generated for simulating the point loading in the experiments. The velocities
% of the top boundary particles are vx = 0,vy = − 2 mm/s, whereas the velocities of the bottom boundary particles are
% as vy = 0. As Johnstone rocks are classified as soft rocks with a relatively small elastic Young's modulus, both artificial
% viscosity and artificial stress are used in this test to stabilise the numerical solutions. The ramp loading is also applied
% to alleviate the dynamic effect due to the sudden loading. The stresses of boundary particles are updated in the same
% way as that in the 3-point bending test. The stresses of real particles near the loading particles are recorded during the
% simulation process, from which the peak load and fracture toughness can be obtained for the comparison with the
% experimental data.
% Figure 10 illustrates the progressive development of fracture and the horizontal velocity profile at the failure stage
% for the semicircular specimen with the rotational angle of 30 . As seen in Figure 10A, some particles near the tip of the
% notch are first damaged, showing that the fracture is initiated at the location of the notch tip. Then, the fracture propa-
% gates towards the loading point (Figure 10B–D), and eventually, a vertical curvilinear fracture pattern is formed in the
% middle of the sample (Figure 10E). The horizontal velocity profile corresponding to Figure 10E is also plotted in
% Figure 10F. It can be seen that the particles on the left side of the fracture move to the left whereas the particles on the
% other side move to the right direction; thus, the sample is split into two halves at the end. In this approach, SPH particle
% itself could not be cracked visually; however, each damaged SPH particle possesses a FPZ, and the material behaviour
% within the FPZ will contribute to the decrease in the overall stress state of the SPH particle, thereby resulting in the sep-
% aration between two damaged SPH particles. The same test without using the artificial stress is also plotted in
% Figure 11. The result remains stable when the fracture is initiated (Figure 11A) but becomes unstable during the post-
% failure process (Figure 11B,C). Unrealistic fracture and particles clustering are observed in Figure 11B,C, which indi-
% cates the tensile instability problem is occurred in SPH. By comparing those two results, it can be concluded that the

% tensile instability problem can be completely removed and good numerical stability can be achieved by just adopting
% this artificial stress term.
% Figure 12 depicts the comparison of mixed-mode fracture envelope between experiments and SPH simulations for
% different notch inclination angles α = 0 , 15  , 30  , 45 , 50 and 60 . It shows that the simulation fracture patterns
% marked as red lines match well with the experimental results shown as black lines. The final fracture pattern is a verti-
% cal straight line when α = 0 (pure mode I). When α > 0 , the sample is subjected to mixed-mode or pure mode II load-
% ing, resulting in a curvilinear fracture paths. For all simulation cases, the fractures initiate from the notch tip and then
% propagate towards upper loading point, while the experiments show that a few fractures tend to initiate behind the
% notch tip when the notch angle is beyond α =50 . Nevertheless, the simulation results still could successfully predict
% the overall fracture envelope in experiments and demonstrate the
% capability of the traditional SPH model for capturing

% the fracture development under various complex loading conditions. In the meanwhile, the comparison of peak load
% between experiments and simulations for the notch inclination angles varying from 0 to 60 with an internal incre-
% ment of 5 is plotted in Figure 13A. It is noted here that a notch angle of 54 (instead of 55) is used in the current test
% as the sample is subjected to pure mode II loading condition when α = 54 .95 The SPH simulation can well capture the
% increasing trend of the peak load with the increasing notch angles in the experiment. Additionally, Figure 13B com-
% pares the predictions of fracture toughness in SPH simulation against the experimental results and Lim's theory.95 The
% simulation results perfectly fall within the experimental data ranges even though the fracture toughness data in the
% experiments are quite scattered. Nevertheless, the numerical results agree closely with the empirical fracture toughness
% envelope, which was proposed by Lim et al.95 and represented as a black line Figure 13B. It further confirms that our
% model could accurately predict the mixed-mode fracture behaviour of rocks in semicircular bending test.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    1540    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{./Figure-Semicircular-bending-test}
  \caption{Geometry and boundary condition of the semicircular bending
    test.}
  \label{fig:geometry-Semicircular-bending-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cite{Wang_2020}

\subsection{Drop-weight impact test}
\label{sec:3.3}
\PNA{This section is devoted to proof
the accuracy of the eigensoftening algorithm within the \acrshort{mpm} framework when the behavior
of quasi-brittle materials under dynamic loading is simulated. One of the most interesting examples of this
loading case is the three-point bending test on a concrete notched beam when it is conducted under impact loading. Indeed, this experimental test is appropriate to be reproduced numerically with the explicit solver proposed in this paper since
fracture occurs in a period of milliseconds and waves propagate fast along the high strength
materials involved in the aforementioned test.} Accordingly, small time steps are
required to achieve a numerically stable simulation.\\

An interesting case of this experimental test was reported by Zhang {\it et al.} \cite{Zhang_2009,Zhang_2010a}. As in the aforementioned study,
an impact hammer of 120.6 kg has been employed in the proposed simulations to drop it at an impact
speed of 881 mm/s. The beam dimensions were 100 mm x 100 mm (B x D)
in cross section, and 420 mm in total lenght (L). The initial
notch-depth ratio was approximately 0.5, and the span, S, was fixed at
300 mm during the tests (see Figure
\ref{fig:geometry-drop-weight-impact-test}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./Figure-impact-test}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-drop-weight-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material adopted for the simulation was characterised by Navas
{\it et al.}~\cite{Navas_2017_ES}, being the material properties, such
as the material density, \gls{rho}, the compressive strenght, \gls{fc},
the tensile strenght \gls{ft}, the specific fracture energy, \gls{Gf},
and the elastic modulus \gls{E}, provided in table \ref{tab:concrete-properties}.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented in order to properly capture the impact. Several levels of discretization were required to assess
the objectiveness of the obtained results. The results here exposed
are from a discretization of 30616 material points and 6297 nodes. To
achieve better results, an unestructured grid layout has been adopted
focusing the minimum nodal spacing, 0.47 mm, in the middle of the beam and in
the impact surface of the hammer. For this simulation, \gls{gamma-lme} was
fixed to 6. Once the accuracy of the \acrshort{lme} approximants have been proven
in Section \ref{sec:3.1}, this discretization has been adopted.
%without comparing
%with linear interpolation. Therefore \\

First, the reaction and impact forces are validated against their
experimental counterparts. Since the impact forces applied by the
hammer and the reaction forces at the two supports were experimentally
measured, they are compared with numerical ones in Figure
\ref{fig:Reactions-Forces-impact-test}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-impact-test-Forces-Time}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-Forces-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Note that the general trend of both forces are correctly
captured. Observe that in experimental solution the delay between
impact peak and reactions peak does not agree with the numerical
ones. On the other hand, if the delay is computed analytically through
concrete wave speed is much closer to the numerical results. A
feasible explanation to this unexpected result can due the composite
nature of the concrete. Another discrepancy is in the much higher
values of maximum impact and reaction forces in numerical than in
experimental. In this case, this is a consequence of over-rigidity
in the supports. \PNA{Although the traditional \acrshort{mpm} implementation allows collision of two bodies, inaccurate energy conservation is obtained without the implementation of a conservative contact algorithm to reproduce the hammer impact in the beam. %In the
%%\acrshort{mpm}, the implementation of this feature can be bypassed to obtain preliminary results. Nevertheless, this kind of algorithm could
%be implemented to get ever more accurate solutions. 
Despite the observed discrepancies, this is
out of the scope of this document, although future research of authors will overcome this
limitation. Similarly, results of support forces will be more realistic when modelled by a suitable contact algorithm instead of avoiding any vertical displacement}.\\

Finally, longitudinal stress distribution and crack propagation are
depicted in Figure \ref{fig:Stress-vs-damage-impact-test}. Stress
evolution agrees with the results obtained by \cite{Navas_2017_ES}
under the \acrshort{otm} framework. Two additional considerations about
this results are concerning the interpolation technique. On the one hand,
in the opposite to the \acrshort{ugimp} shape function, \acrshort{lme}
allows unstructured grids. This is a remarkable benefit for this kind
of simulations since the majority of the computational effort is
clearly localised in a tiny region of the domain. On the other
hand, there is the possibility to reduce the shape function support to avoid
a material point in one side of the crack is affected by those in the
other side of the crack. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 0.27 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0060}}
\subfigure[t = 0.36 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0080}}
\subfigure[t = 0.45 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0100}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-vs-damage-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:4}
\PNA{Based on the EigenMPM approach to fracture proposed by Zhang {\it et al.}
\cite{Zhang_EE_2020}, two additional improvements have been proposed in the present research in order to 
solve the shortcomings of the original concept that Zhang {\it et al.} pointed out in the aforementioned paper \cite{Zhang_EE_2020}.

Firstly, by
introducing the concept of eigensoftening to the \acrshort{mpm} framework,
which allows to consider gradual failure process of each material point. This concept is analogous the softening law which is traditionally employed in the context
of cohesive fracture in \acrfull{fem}. Furthermore, this failure criterion admits the design of more elaborated damage curves, which can be employed in order to describe the failure process in more sophisticated
materials. In this direction, an interesting future improvement of the
algorithm would be the ability to simulate of crack patterns in
composite material such as steel or carbon fiber reinforced concretes\cite{Ruiz2019}. Interesting researches have been done on this field~\cite{Navas_2018_ES}. In addition, the eigensoftening approach can be extrapolated to rock fracture in the same manner the eigenerosion was~\cite{Wang2017}, taking into account that the volumetric stiffness remains in order to support bulk loads.

Secondly, \acrshort{lme} is employed to remove from
the solution those wiggles provoked by inaccuracies in the stress integration. This spatial discretization was previously employed within the \acrshort{mpm} framework with excellent results under the elastic regime~\cite{Wobbes2020,Molinos2020}. The traditional time scheme of the \acrshort{mpm} has been improved with a Newmark Predictor Corrector scheme as well. In this research, the aforementioned techniques have helped to improve the
results in fracture propagation through the eigensoftening and eigenerosion algorithms. Further improvement may be reached by adapting \acrshort{lme} through the
deformation gradient \cite{Kochmann2019}. \textit{A priori} this would enhance the localisation properties of the proposed algorithms in more sophisticated domains.

Finally, for better capturing the measurement of forces made in the
laboratory, more realistic contact tools have to be
developed. Navas~\textif{et al.} demonstrate the importance of the
stiffness of the support in order to obtain accurate results of the
reaction forces. Interesting contact algorithms within the
\acrshort{mpm} methodology are available in the literature
\cite{Bardenhagen_Contact_2001,XZhang_Contact_2011}}

@@

\end{document}
\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-1244) }
